---
title: "eds221-day4-interactive"
author: "carly"
format: html
editor: visual
---

## Functions

Make a function that doubles a value.

```{r}
double_value <- function(my_dollars){
  print(my_dollars * 2)
}
double_value(25)
double_value(my_dollars = 25)
```

```{r}
#We can create a vector of values and fun that through our function
monthly_income <- c(42.50, 61.75, 89.50)
double_value(monthly_income)
```

Create a function that returns an awesome restaurant name given a cool animal and your favorite food:

```{r}
name_restaurant <- function(animal, food){
  print(paste0(animal, food))
}

name_restaurant("rabbit", "soup")
```

Given temperature in degrees Fahrenheit, write a function that converts it to Celsius.

```{r}
temp_f_to_celsius <- function(fahrenheit){
  paste(.555 * (fahrenheit - 32))
}
temp_f_to_celsius(52)
```

Create a function that given inputs of age and hometown returns "I am \_years old and I am from \_"

```{r}
age_hometown <- function(age, hometown) {
  print(paste("I am", age, "years old and I am from", hometown))
}
age_hometown(age = 27, hometown = "Shelburne")
```

```{r}
#We also don't have to use print in the function because that is saying "show me this". We can just run the function and return and store the value as seen below. We can then reuse that stored value for calculations or in other work we are doing. 
double_value2 <- function(my_dollars){
  my_dollars * 2
}
double_value2(10) 
```

## Conditions in Functions

```{r}
calculate_animal_age <- function(animal, age_human_years){
  if (animal == "dog"){
    age_human_years * 7
  } else if (animal == "goat"){
    age_human_years * 4.7
  } else {paste("Animal not in logic. Please add dog or goat")}
}
calculate_animal_age("dog", 4)
calculate_animal_age("goat", 5)
calculate_animal_age("bird", 3)
#We can see that if we add a part of the function that is not defined, it does not return anything so we must add an else statement. 
```

Write a function that, given temperature in Kelvin, if a user selects "metric" for the scale then return celsius, or if they select "english" then return degrees in Fahrenheit.

```{r}
temp_conversion <- function(scale, temp_in_K) {
  if (scale == "metric"){
    print(paste((temp_in_K - 273.15), "degrees celsius"))
  } else if (scale == "english") {
   print(paste(((temp_in_K - 273.15) * (9/5) + 32), "degrees fahrenheit"))
  } else {print("Please select metric or english scale.")}
}
temp_conversion("english", 450)
```
